A few stickleback suffice to transport adaptive alleles to new lakes

### 2018-12-11 Stickleback_Paper.tex

\paragraph{Habitat and geography}
%The simulation begins with two subpopulations, one of both the
Throughout the simulation there are two habitat types defined by their selective pressures.
We start the simulations with two populations of size 5,000 diploid individuals in each of the habitats, for a total of 10,000 diploid individuals.
%The two habitat types -- made to model the marine and freshwater selective pressures --  a fixed total populations size of 5,000 diploid individuals each. 
%The habitat types differ in the optimal trait value.
%The simulations begin with two subpopulations, each of size 5,000 diploid individuals, one set living in marine and freshwater selective pressures.
The arrangement of these habitats, depicted in Figure \ref{fig:Geo},
roughly models a set of freshwater habitats along a stretch of coastline. 
The marine habitat is a continuous, one-dimensional range of length 25 units,
while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''),
each connected to the marine habitat at regularly spaced intervals
(positions $i - 1/2$ for $1 \le i \le 25$).

Divergent selection is mediated by a single quantitative trait
with different optima in marine and freshwater habitats.
This situation roughly models the cumulative effect of the various phenotypes such as armor morphology, 
body size, craniofacial variation and opercle shape on which divergent selection is thought to act on in the two environments. 
Concretely, the optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$
is determined by a Gaussian kernel with standard deviation 15, i.e.,
\begin{align*}
    f(x_\text{ind}; x_\text{opt})
    &=
    \exp\left\{
        \frac{1}{2}
            \left(
            \frac{x_\text{ind}-x_\text{opt}}{15}
            \right)^2
        \right\} .
\end{align*}
Note that only the scale of trait values relative to mutation effect size is relevant.
In short, the difference between an individual's trait value and the optimum is the main determinant of an individuals fitness.
We chose the difference between optima and strength of stabilizing selection in each habitat
so that 
(a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,
and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.

@@

\textbf{This next paragraph needs work, skipping for now}
Dispersal of individuals occurs both locally along the coastline in the marine habitat 
and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage (before the selection of parents). 
There is no dispersal directly between, or within, lakes.
The bi-directional lake--ocean migration rate is denoted $m$, 
which we refer to as the rate of \emph{gene flow} between habitats.
Parents for a new lake individual are chosen from the marine habitat with probability $m$:
to obtain the pair of freshwater parents, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake. 
Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$): 
first, a single parent is chosen proportionally to fitness in the marine habitat,
and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
where $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, 
and reflected to stay within the population.
New offspring in the freshwater habitat are chosen in the same way, except the probability that the parents are marine individuals is $m$; 
any new freshwater offspring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.

@@

%%%%%%%%%%%%%%%%%%
\section*{Results}

We varied the ocean--lake migration rate, $m$, 
across separate simulations
from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
Since each lake contains 200 individuals,
this corresponds to between 0.01 and 100 migrants per lake per generation.
Many aspects of adaptation changed substantially across this range,
including the speed of adaptation,
degree of sharing of adaptive alleles between lakes,
and the population genetic signals left behind.
At very low rates of gene flow,
each new lake's population adapted \emph{de novo},
which took a very long time.
At very high rates of gene flow, local adaptation was almost impossible.
Between these two extremes,
genetic variation that allowed adaptation to freshwater habitats
could move relatively easily between lakes --
despite being deleterious in the intervening marine habitat --
which allowed populations arriving in new lakes to adapt quickly,
reassembling collections of alleles responsible for freshwater adaptation
in other lakes.

%%%%%%%%%%%%
\subsection*{Local Adaptation: differentiation with gene flow}

@@

As expected, increasing migration rate decreased differentiation between habitats. 
As seen in Figure~\ref{fig:Fst}, 
$F_{ST}$ between marine and freshwater habitats 
at neutral sites steadily declines as migration increases. 
However, local adaptation was still able to occur despite overall homogenization 
at up to 10 migrants per lake per generation.
% if computed using only sites with alleles affecting the trait (``effect mutations''),
% $F_{ST}$ between habitats was relatively constant across migration values.

@@

\begin{figure}
	\begin{center}
  		\includegraphics{Final_Plots/Time_Adapt.pdf}
  		\caption{
		Time to adaptation as a function of migration rate.
        The time to adaptation is measured as the number of generations until
		the introduced population's mean phenotype 
        comes within 0.5 of the original lakes average phenotype. 
		Each point represents a single simulation run,
		and the yellow dashed line is the average of all points at each respective parameter value.
        (Adaptation did not occur at the highest rate of gene flow.)
        \textbf{where are the multiple points?, They do not exist, we only ran the one set. They each take $\approx$ 36 hours to run; but I can run them again if you want.} 
        } \label{fig:TimeToAdaptation}
	\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Sharing of freshwater adapted alleles}

%FIGURES THAT SHOW THIS:
%-MPAA / IND
At low migration rates, the \emph{initial} period of adaptation
takes roughly ten times longer for lakes as it does for the ocean.
This difference occurs because at low migration rates, 
adaptation occurs independently in each lake,
and the marine habitat has ten times as many individuals,
and therefore ten times the influx of new alleles than any one lake. 
At higher migration rates, greater mixing
allows lakes to share alleles instead of developing their own genetic basis of adaptation. 
As a first indication of this, Figure \ref{fig:Fst} 
shows that $F_{ST}$ between the ``original'' and ``introduced'' sets of lakes 
at effect mutations decreased with migration rate.

@@

We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
Selection is able to rebuild the freshwater haplotype from marine populations as a medium between all freshwater populations. 
Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.

%Systems in which there is one, static selective pressure with a wide range connected to many, much smaller systems

We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
differences between populations that have little gene flow between them.
It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
mistaken for being causative as they often display the same amount of divergence.

>>>>

\paragraph{Habitat and geography}
Throughout the simulation there are two habitat types -- Marine and Freshwater-- defined by their selective pressures.
We start the simulations with two populations of size 5,000 diploid individuals in each of the habitats, for a total of 10,000 diploid individuals.
The arrangement of these habitats, depicted in Figure \ref{fig:Geo},
roughly models a set of freshwater habitats along a stretch of coastline. 
The marine habitat is a continuous, one-dimensional range of length 25 units,
while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''),
each connected to the marine habitat at regularly spaced intervals
(positions $i - 1/2$ for $1 \le i \le 25$).

Divergent selection is mediated by a single quantitative trait
with different optima in marine and freshwater habitats.
This situation roughly models the cumulative effect of the various phenotypes such as armor morphology, 
body size, craniofacial variation and opercle shape on which divergent selection is thought to act on in the two environments. 
Concretely, the optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$
is determined by a Gaussian kernel with standard deviation 15, i.e.,
\begin{align*}
    f(x_\text{ind}; x_\text{opt})
    &=
    \exp\left\{
        \frac{1}{2}
            \left(
            \frac{x_\text{ind}-x_\text{opt}}{15}
            \right)^2
        \right\} .
\end{align*}

In short, the difference between an individual's trait value and the optimum largely determine's that individuals fitness. 
Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
We chose the difference between optima and strength of stabilizing selection in each habitat
so that 
(a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,
and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.

@@

Dispersal occurs both locally along the coastline in the marine habitat 
and between the marine habitat and the lakes, and can be thought of as occurring at the juvenile stage. 
There is no dispersal directly between lakes.
The lake--ocean migration rate is denoted $m$, 
which we refer to as the rate of \emph{gene flow} between habitats.
Each new individual in the marine habitat has freshwater parents with probability $m$; 
to obtain the pair, a first parent is chosen proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. 
The resulting offspring is given a spatial location in the marine habitat at the location of the parent's lake. 
Parents for a new marine individual who is not a migrant are chosen similarly (with probability $1-m$): 
first, a single parent is chosen proportionally to fitness in the marine habitat,
and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
where $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.02, 
and reflected to stay within the population.
New offspring in the freshwater habitat are chosen in the same way, except the probability that the parents are marine individuals is $m$; 
any new freshwater offspring produced by marine individuals are assigned to the lake nearest to the position of the first marine parent.
\textbf{The paragraph above has a confusing structure, skipping for now}

@@

\section*{Results}

To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
Since each lake contains 200 individuals, we refer to these migration rates as between 0.01 and 100 migrants per lake, per generation.
Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
At very low rates of gene flow, each new lake's population adapted completely separately from \emph{de novo} mutation, which took a very long time ( $\approx 20,000$ generations).
At very high rates of gene flow, local adaptation was almost impossible.
Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
Surprisingly, only a few migrant per generation from the lakes to the marine were needed to see the effects of rapid local adaptation accelerated by selection on standing genetic variants in the new lakes.
despite being deleterious in the intervening marine habitat --
which allowed populations arriving in new lakes to adapt quickly, reassembling collections of alleles responsible for freshwater adaptation in other lakes.

\subsection*{Local Adaptation: differentiation with gene flow}

@@

As expected, increasing migration rate decreased differentiation between habitats. 
As seen in Figure~\ref{fig:Fst}, 
$F_{ST}$ between marine and freshwater habitats 
at neutral sites steadily declines as migration increases. 
However, local adaptation was still able to occur despite overall homogenization 
at up to 10 migrants per lake per generation.

@@

\begin{figure}
	\begin{center}
  		\includegraphics{Final_Plots/Time_Adapt.pdf}
  		\caption{
		Time to adaptation as a function of migration rate.
        		The time to adaptation is measured as the number of generations until
		the introduced population's mean phenotype 
        		comes within 0.5 of the original lakes average phenotype. 
		Each point represents a single simulation run.
        		(Adaptation did not occur at the highest rate of gene flow.)
        } \label{fig:TimeToAdaptation}
	\end{center}
\end{figure}

\subsection*{Sharing of freshwater adapted alleles}

At low migration rates, the \emph{initial} period of adaptation
takes roughly ten times longer for lakes as it does for the ocean.
This difference occurs because at low migration rates, 
adaptation occurs independently in each lake,
and the marine habitat has ten times as many individuals,
and therefore ten times the influx of new alleles than any one lake. 
At higher migration rates, greater mixing
allows lakes to share alleles instead of developing their own genetic basis of adaptation. 
As a first indication of this, Figure \ref{fig:Fst} 
shows that $F_{ST}$ between the ``original'' and ``introduced'' sets of lakes 
at effect mutations decreased with migration rate.

@@

We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
Selection is able to rebuild the freshwater haplotype at a rapid pace (in tens of generations) from marine populations as a medium between all freshwater populations. 
Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum. 
Surprisingly, we have discovered the genetic basis for local adaptation which accelerated by standing genetic variation to be derived largely from the initial generation of inhabitants. 
This is in opposition to the assumption that subsequent migration is also necessary for this process to take place.
In turn, this means that the downstream movement of alleles \emph{from} the lakes \emph{to} the oceans is much more important in this system than upstream migration events. 
We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
differences between populations that have little gene flow between them.
It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
mistaken for being causative as they often display the same amount of divergence.
